---
title: "Joining and pivoting datasets"
author: "Your-name-here"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## 2. Pivoting

For this exercise, You will use the following datasets, which live in the `tidyverse` library, and are a bit messy: 

```{r}
table4b <- table4b
table4a <- table4a
```

`table4a` contains counts of cases for 1999 and 2000 across several countries. `table4b` contains population values for 1999 and 2000 across several countries 

#### Q2.1. Have a look at both datasets. 

```{r}

```

#### Q2.2. Which are the columns that should be pivoted longer?

#### Q2.3. Pivot the tables longer, and save them as `table4a` and `table4b` (overwrite the old versions)

```{r}


```


## 3. Joining 

Now, let's plan to join the two datasets that were pivoted longer: `table4a` and `table4b`. 

#### Q3.1. Let's start be identifying the columns that are shared. Visualize the structure of both dataframes, then list the columns that are repeated: 

```{r}

```

#### Q3.2. Given the columns that overlap between the datasets, what kind of join would be appropriate? If you are choosing `left_join()` or `right_join()`, specify which dataframe is on the right or left. 


#### Q3.3. Perform the join, specifying column(s) to join by, and save the resulting dataframe as a new object. Print the new dataframe to check your work.

```{r}


```

