---
title: "Intro to plotting with ggplot2"
author: "Your-name-here"
date: '2023-01-26'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)



```

## Set up 

In this exercise, you will use `ggplot2` which will be loaded when you load `tidyverse`. You will also use the `penguins` dataset we've used before, which lives in the `palmerpenguins` package. 

Look at the code chunk below. The code for loading the `tidyverse` library is already there. Modify the following line so you also load the `palmerpenguins` library. Make sure you feel comfortable with how loading packages works - you'll be expected to do that on your own.

```{r }
library(tidyverse)
---()

```

Now, save the penguins dataset into your environment as `penguins`. That code is already written for you. Once you have loaded the dataframe, look at its structure and/or print the first few rows to get a sense for the dataset.

```{r}
penguins <- penguins 
str(---)
head(---)
```

## 1. Body mass in the penguins

Q1. We want to understand the distribution of `body_mass_g` in this dataset. What type of variable is this?



Q2. Given the type of variable, what type of plot should we use?



Q3. Build the plot - remember to grammar of graphics principles and refer to your notes or readings for support

```{r}
ggplot(penguins, aes(x = ---, y = ---))+
  geom_----()

```

Q3. Let's now modify this plot. You likely saw a warning pop up when running that last command that `ggplot2` automatically used 30 bins. "Bins" are the intervals used to divide up values of your continuous variables into a series of non-overlapping ranges into which individuals can be counted. Copy and paste your last set of code into this new chunk and modify it. Let's set a binwidth of 250, meaning that we would like to have even intervals of 250 units along the x axis.

*Hint*: use your error message, your grammar of graphics concepts, the `?` and help viewers, resources on Brightspace, and the internet if you need support getting this done correctly!

```{r}


```


Q4. Sometimes, it can be a bit difficult to see the differences between bars in a histogram. To help, let's adjust some of the colors in the plot. Copy and paste your last plot, and edit it to have two new arguments in your plotting layer: `color` and `fill`. Set `color` to `black` ; then set `fill` to whichever color you like! You can refer to your resources to see what colors are available to you.

```{r}


```


Q5. What element of your plot did `fill` and `color` modify? 



Q6. Using the plot that you have created, which weight interval has the most individuals? 




## 2. Body mass by species of penguins 

Let's now look at how body mass varies across species of penguins. 

Q1. You already know what type of variable is contained in the `body_mass_g` vector. What type of variable is contained in the `species` vector? 



Q2. Armed with the knowledge that these are the two variables you want to visualize, which type of plot should you use? 



Q3. Create this to plot. The skeleton of the code you will need is prepared for you below: 

```{r }
ggplot(---, aes(x= ---, y = ---))+
  geom_---()

```


Q2. Which species of penguins is the heaviest?



Q3. Let's enhance the communicative value of our plot. Copy and paste the code for the original plot, then modify it. How would you color the boxes by species? 

```{r}


```



Q4. Modify the plot again - how would you add a title and axis labels? 

```{r}


```


Q5. Add another layer to the plot, this time a violin layer (`geom_violin()`). Consider the grammar of graphics - each layer is added on top of the last. Where in the line of code would you add it if you still wanted the boxplot to be visible?  

```{r}


```


Q6. Copy and paste the code from Q4, and modify the plot a bit to make it prettier. Play with the `width` argument for the boxplot layer so that it fits inside the violin distribution. 

```{r}


```


Q7. Lastly, we now have both axis labels on the x axis and a legend of colors to the right of our plot. This information is redundant, and we can increase the size of our plot by removing the legend. Add a layer to your previous plot that removes the legend from your plot.

*Hint*: This task is a bit harder, and it requires you to search for the answer in the resources listed on Brightspace and/or by searching for help on the internet. Try a few solutions before asking for help - troubleshooting is an important skill!

```{r}


```

